---
title: "PA1_template.Rmd"
author: "Alexander Johnston"
date: "June 30, 2017"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Loading and preprocessing the data
```{r}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activity<- read.csv('activity.csv')
```

### What is the mean total number of steps taken per day

Made a histogram of the total number of steps taken each day, with the following code:
```{r}
dailystep <- tapply(activity$steps, activity$date, FUN = sum, na.rm = TRUE)
hist(dailystep, breaks=20, col="red")
```

Working out the mean and median average:
```{r}
stepmean <- mean(dailystep)
stepmedian <- median(dailystep)
```

The `mean` is `r stepmean` and the `median` is `r stepmedian`.

### What is the average daily activity pattern?

Made a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
stepsbyinterval <- aggregate(steps ~ interval, activity, mean)
plot(stepsbyinterval$interval,stepsbyinterval$steps, type="l", main="Average Daily Step Numbers by Interval", xlab="Interval", ylab="Step Number", col="red")

interval_largest <- stepsbyinterval[which.max(stepsbyinterval$steps),1]
```

The 5-minute interval, on average across all the days in the dataset, which has the maximum number of steps is `r interval_largest`.

### Imputing missing values

Calculated the total number of missing values in the dataset

```{r}
MissingValues <- length(which(is.na(activity$steps)))
```

Created a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
library(Hmisc)
Imputedactivity <- activity
Imputedactivity$steps <- impute(Imputedactivity$steps, fun=mean)
```


Made a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. 

```{r}
dailystep2 <- tapply(Imputedactivity$steps, activity$date, sum, na.rm=TRUE)
hist(dailystep2, col= "blue", breaks=seq(from=0, to=25000, by=1500), main="Total Number of Steps Taken Each Day", xlab="Total Number of Steps")
```

Working out the mean and median average:
```{r}
stepmean2 <- mean(dailystep2)
stepmedian2 <- median(dailystep2)
```

The `mean` is `r stepmean2` and the `median` is `r stepmedian2`.

Mean and median values are higher after inputing the missing data.


### Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
activity$dateType <-  ifelse(as.POSIXlt(activity$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
avgactivity <- aggregate(steps ~ interval + dateType, data=activity, mean)
ggplot(avgactivity, aes(interval, steps, color="Red")) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("average number of steps")
```
